Unbalanced Mallows Models for Optimizing Expensive Black-Box Permutation Problems

### 2021-02-04 paper/main.tex
paper s5

\subsection{Rankings and orderings}
In Section~\ref{sec:preliminaries} we highlighted the great difference of rankings and orderings, usually ommitted in the literature. In this lines, we show the consequences of misinterpreting permutations. We remind the reader that statistical models (Borda or the \ken distance) are defined for rankings while standard literature in optimization considers orderings. This is the reason while in the optimization field we will usually find the Swap distance instead of the \ken distance. However, both are usually claimed to be the same, which is false: they are the same in particular cases, for example when  one of the permutations is the identity or an inversion.

@@

\subsection{Analysis of UMM and CEGO on LOP and PFSP}
To  understand the behavior of the algorithms, we record the fitness of
the solution evaluated at each step of each run and we average those values
over 10 runs with different random seeds on the same problem instance.  As an
example, each line in Fig.~\ref{fig:results} shows the mean fitness of the
solutions evaluated by either UMM or CEGO on one instance. The
shaded area shows one standard deviation around the mean. Ideally, each new
evaluation will monotonically improve in fitness, since each solution evaluated
helps to refine the model and leads to a better solution being evaluated in the
next step.  However, it is possible that new solutions are worse
than their predecessors. %Nevertheless, in this example, there is a clear
% overall improvement as more evaluations are performed.
The minimum values
achieved in the plot up to a particular number of function evaluations gives an
estimation of the mean fitness of the best-so-far solution up to that point in
the run. The algorithms do not return the final solution shown but rather the
best one (minimum fitness) of each run. This result summarization is not the classical. However, due to the limited number of function evaluations, we can have a better picture of the behaviour of the algorithm at each step. As a consequence, results displayed here are not comparable with the original paper of CEGO: for this comparison we provide in Table~\ref{table:fitness} the summary of the best solutions found by each algorithm.

Figure~\ref{fig:lolib} shows the results on the real-world instances taken from LOLIB. 
  %% MANUEL: Ya dimos la URL arriba, así que no la repetiría aquí
  % \footnote{\url{http://grafo.etsii.urjc.es/optsicom/lolib/}}
%The type of instances differs greatly from the synthetic LOP instances, and,
%thus, so does the landscape. However, UMM shows a very similar behavior: 
We can see that for UMM there 
is a (close to) linear decrease of the fitness of new solutions over each new
evaluation. This suggests that UMM is able to model the fitness
landscape. Regarding CEGO, \ekhine{we observe a similar behavior to UMM for the first
50-100 evaluations, however, in several instances CEGO reaches the time limit
of 144 hours, so no observations can be made for longer budgets. In smaller
instances (e.g., Figure~\ref{fig:lolib}, top), CEGO again appears to stagnate.}{ the objective function decrease slowly?}

>>>>

\subsection{Rankings and orderings: experiments}\label{sec:exper_ro}
\ekhine{}{este primer parago es bastante importante para dejarlo aqui. }
In Section~\ref{sec:preliminaries} we highlighted the great difference of rankings and orderings, usually ommitted in the literature. In this lines, we show the consequences of misinterpreting permutations. We remind the reader that statistical models (Borda or the \ken distance) are defined for rankings while standard literature in optimization considers orderings. This is the reason while in the optimization field we will usually find the Swap distance instead of the \ken distance. However, both are usually claimed to be the same, which is false: they are the same in particular cases, for example when  one of the permutations is the identity or an inversion.

@@

\subsection{Analysis of UMM and CEGO on LOP and PFSP}
To  understand the behavior of the algorithms, we record the fitness of
the solution evaluated at each step of each run and we average those values
over 10 runs with different random seeds on the same problem instance.  As an
example, each line in Fig.~\ref{fig:results} shows the mean fitness of the
solutions evaluated by either UMM or CEGO on one instance. The
shaded area shows one standard deviation around the mean. Ideally, each new
evaluation will monotonically improve in fitness, since each solution evaluated
helps to refine the model and leads to a better solution being evaluated in the
next step.  However, it is possible that new solutions are worse
than their predecessors. %Nevertheless, in this example, there is a clear
% overall improvement as more evaluations are performed.
The minimum values
achieved in the plot up to a particular number of function evaluations gives an
estimation of the mean fitness of the best-so-far solution up to that point in
the run. The algorithms do not return the final solution shown but rather the
best one (minimum fitness) of each run. This result summarization is not the classical. However, due to the limited number of function evaluations, we can have a better picture of the behaviour of the algorithm at each step. As a consequence, results displayed in the figures are not comparable with the original paper of CEGO: for this comparison we provide in Table~\ref{table:fitness} the summary of the best solutions found by each algorithm, which are similar to the original study.

\paragraph{Figure descriptiopn} Figure~\ref{fig:results} shows the results on the real-world instances from LOLIB and PFSP. There is one figure per instance. In each figure there is one horizontal line displaying the best known solutions for the problem. Besides, there are 5 different lines representing each of the different algorithm/versions considered, in particular:
\begin{itemize}
\item UMM\_100, UMM\_200 and UMM\_400 are three different versions for UMM, each considering a different total number of function evaluations, 100, 200 and 400. Obviously, the lines corresponding to UMM\_100, UMM\_200 wont reach the right corner of the figure. 
\item CEGO\_orig and CEGO\_inv are, respectively, the original version of CEGO and the modification of Section~\ref{sec:exper_ro}, i.e., the ranking generated by the CEGO algorithm at each iteration, $\sigma$, is inverted before being evaluated, so the evaluation is performed on $\sigma^{-1}$. 
\end{itemize}

We start our analysis with the three version of UMM. We can see a general trend for all the instances: After the 10 random initial permutations, the fitness of the new solutions decreases with a linear trend along iterations. The (close to) linear decrease is given by the linear decrease in the expected distance of the model which is sampled at each step, described in Section~\ref{sec:sampling}. Overall, the improvement does not seem to converge not even for the longer runs of 400 evaluations.
The average runtime for all the instances is around 0.3 minutes, 1.15 minutes and 4.5 minutes respectively for the  versions with 100, 200 and 400 fitness evaluations.

  %% MANUEL: Ya dimos la URL arriba, así que no la repetiría aquí
  % \footnote{\url{http://grafo.etsii.urjc.es/optsicom/lolib/}}
%The type of instances differs greatly from the synthetic LOP instances, and,
%thus, so does the landscape. However, UMM shows a very similar behavior: 
We can see that for UMM there 
is a (close to) linear decrease of the fitness of new solutions over each new
evaluation. This suggests that UMM is able to model the fitness
landscape. Regarding CEGO, \ekhine{we observe a similar behavior to UMM for the first
50-100 evaluations, however, in several instances CEGO reaches the time limit
of 144 hours, so no observations can be made for longer budgets. In smaller
instances (e.g., Figure~\ref{fig:lolib}, top), CEGO again appears to stagnate.}{ the objective function decrease slowly?}
